---
title: "Machine Learning Project"
author: "Ivan Tsatsarov"
date: "April 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data Load

```{r load}
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")

```

#Data Clean
Remove Variables that have more than 50 % of NA.
Remove variables with zero variance
Remove first identity column
Remove timestamp columns

```{r clean}
library(caret)
nz<-nearZeroVar(training, saveMetrics = TRUE)
training<-training[,!nz$nzv]
more50NA<-function(x) 
  if(sum(is.na(training[, x])) > 0.50*nrow(training))    {return(TRUE)
}else{
return(FALSE)
}
c50NA<-sapply(colnames(training),more50NA)
training<-training[,!c50NA]
training<-training[,c(2:59)]
training<-training[,-c(2,3,4)]

```


## Cross Validation sets

From Training data I separate 70 % as training and 30 %as test

```{r cross}
library(caret)
set.seed(1234)
my_train <- createDataPartition(y=training$classe,p=.70,list=F)
my_training <- training[my_train,]
my_testing <- training[-my_train,]
```

##Testing different models
```{r model eval:FALSE,cache:TRUE,results:"hide"}
tc <- trainControl(method = "cv", number = 1, verboseIter=FALSE , preProcOptions="pca", allowParallel=FALSE)

rf <- train(classe ~ ., data = my_training, method = "rf",tControl=tc)
svml <- train(classe ~ ., data = my_training, method = "svmLinear",tControl=tc)

ldam <- train(classe ~ ., data = my_training, method = "lda",tControl=tc)

rdam <- train(classe ~ ., data = my_training, method = "rda",tControl=tc)

rf$results$Accuracy
svml$result$Accuracy
ldam$results$Accuracy
rdam$results$Accuracy
```

# Accuracy for 4 methods is 
Random Forest - 0.9953
Linear Support Vector Machine - 0.80
Linear Discriminant Analysis -0.74
Regularized Discriminant Analysis -0.91

From Accuracy values I decide to use Random Forest
```{r predictions cache:TRUE,eval=FALSE}
predict_rf<-predict(rf,testing)
predict_matrix<-data.frame(testing$X,predict_rf)
```
